\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-06 Documentation/TearingModeTheory.tex
Added more to tearing mode writeup

\documentclass[notitlepage,12pt]{article}
\usepackage {amsmath}
\pdfoutput = 1 
\usepackage {graphicx}
\usepackage{fullpage}
\allowdisplaybreaks
\newcommand {\bpsi} {\mbox{\boldmath$\psi$}}
\newcommand {\brho} {\mbox{\boldmath$\rho$}}
\newcommand{\bmu}{\mbox{\boldmath$\mu$}}
\newcommand{\bzeta}{\mbox{\boldmath$\zeta$}}
\newcommand{\bpi}{\mbox{\boldmath$\pi$}}
\newcommand{\bsigma}{\mbox{\boldmath$\sigma$}}
\newcommand{\bomega}{\mbox{\boldmath$\omega$}}

@@

\subsection{Fundamental Quantities}
Consider an idealized tokamak plasma consisting of an equal number of electrons, with
mass $m_e$ and charge $-e$ (here, $e$ denotes the  magnitude of the electron
charge), and ions, with  mass $m_i$ and charge $+e$. We shall employ 
the symbol
\begin{equation}
T_s = \frac{1}{3} \,m_s\, \langle v_s^{2}\rangle
\end{equation}
to denote a {\em kinetic temperature}\/   measured in units of energy.
Here, $v$ is a particle speed, and the angular brackets denote an
ensemble average (Reif 1965). The kinetic temperature of species $s$ is a measure of 
the mean kinetic energy of particles of
that species. (Here, $s$ represents either $e$ for electrons, or $i$ for ions.) In tokamak plasma physics, kinetic temperature
is invariably measured in {\em electron-volts}\/ \index{electron-volt} (1 joule is equivalent to
$6.24\times 10^{18}$\,eV).

@@

\subsection{Braginskii Equations}\label{sbrag}
The electron and ion fluid equations in a collisional plasma
take the form:
\begin{align}\label{e3.79a}
\frac{d_en}{dt} + n\,\nabla\cdot{\bf V}_e &= 0,\\[0.5ex]
m_e \,n\,\frac{d_e {\bf V}_e}{dt} + \nabla p_e+ \nabla\cdot \bpi_e + e\, n\,
({\bf E} + {\bf V}_e\times {\bf B})& = {\bf F},\label{e3.79b}\\[0.5ex]
\frac{3}{2}\frac{d_e p_e}{dt} + \frac{5}{2}\,p_e\,\nabla\cdot{\bf V}_e
+ \bpi_e:\nabla{\bf V}_e+ \nabla\cdot{\bf q}_e &= w_e,\label{e3.79c}
\end{align}
and
\begin{align}\label{e3.80a}
\frac{d_i n}{dt} + n\,\nabla\cdot{\bf V}_i &= 0,\\[0.5ex]
m_i \,n\,\frac{d_i {\bf V}_i}{dt} + \nabla p_i + \nabla\cdot \bpi_i - e\, n\,
({\bf E} + {\bf V}_i\times {\bf B})& =- {\bf F},\label{e3.80b}\\[0.5ex]
\frac{3}{2}\frac{d_i p_i}{dt} + \frac{5}{2}\,p_i\,\nabla\cdot{\bf V}_i
+ \bpi_i:\nabla{\bf V}_i+ \nabla\cdot{\bf q}_i &= w_i,\label{e3.80c}
\end{align}
respectively. 
Equations (\ref{e3.79a})--(\ref{e3.80c})  are called the {\em Braginskii equations}, because they were first obtained
in a celebrated article by S.I.~Braginskii (Braginskii 1965). Here, $n$ is the electron number density, ${\bf V}_s$ is the species-$s$ flow velocity, $p_s=n\,T_s$ is the 
species-$s$ scalar
pressure, $\bpi_s$ is the species-$s$ viscosity tensor, ${\bf F}$ is the collisional friction force density, ${\bf q}_s$ is the species-$s$ heat flux density, $w_s$ is the species-$s$ collisional heating rate density, and ${\bf E}$ is the ambient electric field-strength. Moreover, $d_e/dt\equiv \partial/\partial t + {\bf V}_e\cdot\nabla$ in the electron fluid equations, (\ref{e3.79a})--(\ref{e3.79c}), whereas  $d_i/dt\equiv \partial/\partial t + {\bf V}_i\cdot\nabla$ in the ion fluid equations,  (\ref{e3.80a})--(\ref{e3.80c}).

A tokamak plasma is highly magnetized. In other words, 
\begin{equation}
{\mit\Omega}_i\,\tau_i, \,\,\,{\mit\Omega}_e\, \tau_e \gg 1,
\end{equation}
 which implies that the electron and ion gyroradii are much  smaller than the
corresponding mean-free-paths between $90^\circ$ collisional scattering events. In this limit, a standard two-Laguerre-polynomial
Chapman-Enskog closure scheme (Chapman \& Cowling 1953) yields
\begin{align}\label{e3.91a}
{\bf F}&= n\,e\left(\frac{{\bf j}_\parallel}{\sigma_\parallel}
+\frac{{\bf j}_\perp}{\sigma_\perp}\right) -0.71\,n\,\nabla_\parallel T_e
-\frac{3\,n}{2\,{\mit\Omega}_e\,\tau_e}\,{\bf b}\times\nabla_\perp T_e,
\\[0.5ex]
w_i &= \frac{3\,m_e}{m_i} \frac{n\,(T_e-T_i)}{\tau_e},\\[0.5ex]
w_e &= -w_i + \frac{ {\bf j}\cdot {\bf F} }{n \,e}.
\end{align}
Here, ${\bf b}={\bf B}/B$ is a unit vector parallel to the magnetic field, and ${\bf j}= n\,e\,({\bf V}_i-{\bf V}_e)$ is the net plasma current density. 
Moreover, the {\em parallel electrical conductivity}, $\sigma_\parallel$,  is given by
\begin{equation}\label{e3.83}
\sigma_\parallel = 1.96\,\frac{n \,e^2\,\tau_e}{m_e}.
\end{equation} 
whereas the {\em perpendicular electrical conductivity}, $\sigma_\perp$,  takes the form
\begin{equation}\label{e3.92}
\sigma_\perp = 0.51\,\sigma_\parallel = \frac{n\,e^2\,\tau_e}{m_e}.
\end{equation}
Note that $\nabla_\parallel(\cdots) \equiv [{\bf b}\cdot\nabla
(\cdots)]\,{\bf b}$ denotes a
gradient parallel to the magnetic field, whereas $\nabla_\perp \equiv
 \nabla-\nabla_\parallel$ denotes a gradient perpendicular to the magnetic
field. Likewise, ${\bf j}_\parallel \equiv ({\bf b}\cdot{\bf j})\,{\bf b}$
represents the component of the plasma current flowing parallel to the
magnetic field, whereas ${\bf j}_\perp \equiv {\bf j} - {\bf j}_\parallel$
represents the perpendicular component of the plasma current.

In a highly magnetized plasma, the electron and ion heat flux densities are written 
\begin{align}
{\bf q}_e &= -\kappa_\parallel^e\,\nabla_\parallel T_e -\kappa_\perp^e\,
\nabla_\perp T_e
-\kappa_\times^e\,{\bf b}\times\nabla_\perp T_e- 0.71\,\frac{T_e}{e}\,{\bf j}_\parallel-
\frac{3\,T_e}{2\,{\mit\Omega}_e\,\tau_e\,e}\,{\bf b}\times{\bf j}_\perp,\label{e3.94a}\\[0.5ex]
{\bf q}_i &= -\kappa_\parallel^i\,\nabla_\parallel T_i -\kappa_\perp^i\,
\nabla_\perp T_i
+\kappa_\times^i\,{\bf b}\times\nabla_\perp T_i,\label{e3.94b}
\end{align}
respectively. Here, the {\em parallel thermal conductivities}, which control the diffusion of heat parallel to magnetic field-lines, are 
given by
\begin{align}\label{e3.87a}
\kappa_\parallel^e &= 3.2\,\,\frac{n\,\tau_e\,T_e}{m_e},\\[0.5ex]
\kappa_\parallel^i &= 3.9\,\,\frac{n\,\tau_i\,T_i}{m_i},\label{e3.87b}
\end{align} 
whereas the {\em perpendicular thermal conductivities}, which control the diffusion of heat perpendicular to magnetic flux-surfaces, take the form
\begin{align}
\kappa_\perp^e &= 4.7\,\frac{n\,T_e}{m_e\,{\mit\Omega}_e^{\,2}\,\tau_e},\\[0.5ex]
\kappa_\perp^i &= 2\, \frac{n\,T_i}{m_i\,{\mit\Omega}_i^{\,2}\,\tau_i}.
\end{align}
Finally, the {\em cross thermal conductivities}, which control the flow of heat within magnetic flux-surfaces,  are written
\begin{align}
\kappa_\times^e &= \frac{5\,n\,T_e}{2\,m_e\,{\mit\Omega}_e},\\[0.5ex]
\kappa_\times^i &=\frac{5\,n\,T_i}{2\,m_i\,{\mit\Omega}_i}.
\end{align}

@@

The tensor $\bpi_0$ describes what is known as {\em parallel viscosity}.
This is a viscosity that controls the variation along magnetic field-lines of the
velocity component parallel to  field-lines.
 The parallel
viscosity coefficients are given by 
\begin{align}\label{e3.89a}
\eta_0^e &= 0.73\,n\,\tau_e\,T_e,\\[0.5ex]
\eta_0^i &= 0.96\,n\,\tau_i\,T_i,\label{e3.89b}
\end{align}

The tensors $\bpi_1$ and $\bpi_2$ describe what is known
as {\em perpendicular viscosity}. This is a viscosity
that controls the variation perpendicular to magnetic field-lines
of the velocity components perpendicular to field-lines. The perpendicular
viscosity coefficients are given by
\begin{align}
\eta_1^e &= 0.51\, \frac{n\,T_e}{{\mit\Omega}_e^{\,2}\,\tau_e},\\[0.5ex]
\eta_1^i &= \frac{3\, n\,T_i}{10\,{\mit\Omega}_i^{\,2}\,\tau_i}.
\end{align}

Finally, the tensors $\bpi_3$ and $\bpi_4$ describe what is known
as {\em gyroviscosity}. This is not really viscosity at all, because the
associated viscous stresses are always perpendicular to the velocity, implying that
there is no dissipation ({\rm i.e.}, viscous heating) associated with
this effect. The gyroviscosity coefficients are given by 
\begin{align}
\eta_3^e &= -\frac{n\,T_e}{2\,{\mit\Omega}_e} ,\\[0.5ex]
\eta_3^i &= \frac{n\,T_i}{2\,{\mit\Omega}_i}.
\end{align}

\subsection{Normalization of Braginskii Equations}\label{s3.9}
As we have just seen, the Braginskii equations contain terms that describe a
very wide range of different physical phenomena. For this reason, they are
extremely complicated. Fortunately, however, it is not generally 
necessary to retain all of
the terms in these equations when investigating tearing mode dynamics in tokamak plasmas. In this section, we shall attempt to construct a systematic
normalization scheme for the Braginskii equations that will, hopefully, enable us
to determine which terms to keep, and which to discard, when investigating
tearing mode dynamics.

 It is convenient to split the friction force density
${\bf F}$ into a component ${\bf F}_U$ corresponding to resistivity, and a
component ${\bf F}_T$ corresponding to the so-called {\em thermal force}. Thus,
\begin{equation}
{\bf F} = {\bf F}_U+{\bf F}_T,
\end{equation}
where
\begin{align}
{\bf F}_U&= n\,e\left(\frac{{\bf j}_\parallel}{\sigma_\parallel}
+\frac{{\bf j}_\perp}{\sigma_\perp}\right), \\[0.5ex]
{\bf F}_T &= - 0.71\,n\,\nabla_\parallel T_e
-\frac{3\,n}{2\,{\mit\Omega}_e\,\tau_e}\,{\bf b}\times\nabla_\perp T_e.
\end{align}
Likewise, the electron collisional energy gain term $w_e$ is split
into a component $-w_i$ corresponding to the energy lost to the ions (in the
ion rest frame), a component $w_U$ corresponding to work done by the friction
force ${\bf F}_U$, and a component $w_T$ corresponding to work done by the
thermal force ${\bf F}_T$. Thus,
\begin{equation}
w_e = -w_i + w_U + w_T,
\end{equation}
where
\begin{align}
w_U &= \frac{{\bf j}\cdot{\bf F}_U}{n\,e},\\[0.5ex]
w_T&= \frac{{\bf j}\cdot{\bf F}_T}{n\,e}.
\end{align}
Finally, it is helpful to split the electron heat flux density ${\bf q}_e$ into
a diffusive component ${\bf q}_{Te}$ and a convective component ${\bf q}_{Ue}$. 
Thus,
\begin{equation}
{\bf q}_e = {\bf q}_{Te} + {\bf q}_{Ue},
\end{equation}
where
\begin{align}
{\bf q}_{Te} &=-\kappa_\parallel^e\,\nabla_\parallel T_e -\kappa_\perp^e\,
\nabla_\perp T_e
-\kappa_\times^e\,{\bf b}\times\nabla_\perp T_e,\\[0.5ex]
{\bf q}_{Ue}&=  0.71\,\frac{T_e}{e}\,{\bf j}_\parallel-
\frac{3\,T_e}{2\,{\mit\Omega}_e\,\tau_e\,e}\,{\bf b}\times{\bf j}_\perp.
\end{align}

@@

The normalization procedure is designed to make all hatted quantities ${\cal O}(1)$.
The normalization of the electric field is chosen 
 such that the ${\bf E}\times{\bf B}$
velocity is of  similar magnitude to the electron fluid velocity. Note that the parallel viscosity
makes an ${\cal O}(1)$ contribution to $\widehat{\bpi}_e$, whereas the gyroviscosity
makes an ${\cal O}(\zeta_e)$ contribution, and the perpendicular viscosity only 
makes an ${\cal O}(\zeta_e^{2})$ contribution. Likewise, the parallel thermal
conductivity 
makes an ${\cal O}(1)$ contribution to $\widehat{\bf q}_{Te}$, whereas the cross
conductivity 
makes an ${\cal O}(\zeta_e)$ contribution, and the perpendicular conductivity only 
makes an ${\cal O}(\zeta_e^{2})$ contribution. Similarly, the parallel components
of ${\bf F}_T$ and ${\bf q}_{Ue}$ are ${\cal O}(1)$, whereas the perpendicular
components are ${\cal O}(\zeta_e)$.

@@

As before, the normalization procedure is designed to make all hatted quantities ${\cal O}(1)$.
The normalization of the electric field is chosen 
 such that the ${\bf E}\times{\bf B}$
velocity is of similar  magnitude to the ion fluid velocity. Note that the parallel viscosity
makes an ${\cal O}(1)$ contribution to $\widehat{\bpi}_i$, whereas the gyroviscosity
makes an ${\cal O}(\zeta_i)$ contribution, and the perpendicular viscosity only 
makes an ${\cal O}(\zeta_i^{2})$ contribution. Likewise, the parallel thermal
conductivity 
makes an ${\cal O}(1)$ contribution to $\widehat{\bf q}_{i}$, whereas the cross
conductivity 
makes an ${\cal O}(\zeta_i)$ contribution, and the perpendicular conductivity only 
makes an ${\cal O}(\zeta_i^{2})$ contribution. Similarly, the parallel component
of ${\bf F}_T$ is  ${\cal O}(1)$, whereas the perpendicular
component is  ${\cal O}(\zeta_i\,\mu)$.

@@

\section{Cylindrical Tearing Mode Theory}
\subsection{Introduction}
The aim of this section is to describe the simplest theory of tearing mode dynamics in tokamak plasmas, according to which the plasma
equilibrium is approximated as a periodic cylinder.

@@

\subsection{Perturbed Magnetic Field}
Consider a tearing mode perturbation that has $m$ periods in the poloidal direction,
and $n$ periods in the toroidal direction, where $m/n\sim {\cal O}(1)$.  We can write the perturbed magnetic field in the form
\begin{equation}\label{e93}
\delta{\bf B} \simeq \nabla\times (\psi\,{\bf e}_z),
\end{equation}
where
\begin{equation}\label{e93a}
\psi(r,\theta,\varphi,t)= \psi(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\varphi)].
\end{equation}
Here, $\varphi=z/R_0$ is a simulated toroidal angle. The perturbed current density becomes
\begin{equation}\label{e94}
\mu_0\,\delta {\bf j} \simeq - \nabla^2\psi\,{\bf e}_z.
\end{equation}
It is assumed that $|\delta{\bf B}|\ll |{\bf B}|$.

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}\simeq 0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

>>>>

\documentclass[notitlepage,12pt]{article}
\usepackage {amsmath}
\pdfoutput = 1 
\usepackage {graphicx}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\allowdisplaybreaks
\newcommand {\bpsi} {\mbox{\boldmath$\psi$}}
\newcommand {\brho} {\mbox{\boldmath$\rho$}}
\newcommand{\bmu}{\mbox{\boldmath$\mu$}}
\newcommand{\bzeta}{\mbox{\boldmath$\zeta$}}
\newcommand{\bpi}{\mbox{\boldmath$\pi$}}
\newcommand{\bsigma}{\mbox{\boldmath$\sigma$}}
\newcommand{\bomega}{\mbox{\boldmath$\omega$}}

@@

\subsection{Fundamental Quantities}
Consider an idealized tokamak plasma consisting of an equal number of electrons, with
mass $m_e$ and charge $-e$ (here, $e$ denotes the  magnitude of the electron
charge), and ions, with  mass $m_i$ and charge $+e$. We shall employ 
the symbol
\begin{equation}
T_s = \frac{1}{3} \,m_s\, \langle v_s^{2}\rangle
\end{equation}
to denote a {\em kinetic temperature}\/   measured in units of energy.
Here, $v$ is a particle speed, and the angular brackets denote an
ensemble average (Reif 1965). The kinetic temperature of species $s$ is a measure of 
the mean kinetic energy of particles of
that species. (Here, $s$ represents either $e$ for electrons, or $i$ for ions.)

@@

\subsection{Braginskii Equations}\label{sbrag}
The electron and ion fluid equations in a collisional plasma
take the form:
\begin{align}\label{e3.79a}
\frac{d_en}{dt} + n\,\nabla\cdot{\bf V}_e &= 0,\\[0.5ex]
m_e \,n\,\frac{d_e {\bf V}_e}{dt} + \nabla p_e+ \nabla\cdot \bpi_e + e\, n\,
({\bf E} + {\bf V}_e\times {\bf B})& = {\bf F},\label{e3.79b}\\[0.5ex]
\frac{3}{2}\frac{d_e p_e}{dt} + \frac{5}{2}\,p_e\,\nabla\cdot{\bf V}_e
+ \bpi_e:\nabla{\bf V}_e+ \nabla\cdot{\bf q}_e &= w_e,\label{e3.79c}
\end{align}
and
\begin{align}\label{e3.80a}
\frac{d_i n}{dt} + n\,\nabla\cdot{\bf V}_i &= 0,\\[0.5ex]
m_i \,n\,\frac{d_i {\bf V}_i}{dt} + \nabla p_i + \nabla\cdot \bpi_i - e\, n\,
({\bf E} + {\bf V}_i\times {\bf B})& =- {\bf F},\label{e3.80b}\\[0.5ex]
\frac{3}{2}\frac{d_i p_i}{dt} + \frac{5}{2}\,p_i\,\nabla\cdot{\bf V}_i
+ \bpi_i:\nabla{\bf V}_i+ \nabla\cdot{\bf q}_i &= w_i,\label{e3.80c}
\end{align}
respectively. 
Equations (\ref{e3.79a})--(\ref{e3.80c})  are called the {\em Braginskii equations}, because they were first obtained
in a celebrated article by S.I.~Braginskii (Braginskii 1965). Here, $n$ is the electron number density, ${\bf V}_s$ the species-$s$ flow velocity, $p_s=n\,T_s$ the 
species-$s$ scalar
pressure, $\bpi_s$ the species-$s$ viscosity tensor, ${\bf F}$ the collisional friction force density, ${\bf q}_s$  the species-$s$ heat flux density, $w_s$ the species-$s$ collisional heating rate density, and ${\bf E}$  the ambient electric field-strength. Moreover, $d_e/dt\equiv \partial/\partial t + {\bf V}_e\cdot\nabla$ and   $d_i/dt\equiv \partial/\partial t + {\bf V}_i\cdot\nabla$.

A tokamak plasma is {\em highly magnetized}. In other words, 
\begin{equation}
{\mit\Omega}_i\,\tau_i, \,\,\,{\mit\Omega}_e\, \tau_e \gg 1,
\end{equation}
 which implies that the electron and ion gyroradii are much  smaller than the
corresponding mean-free-paths between $90^\circ$ collisional scattering events. In this limit, a standard two-Laguerre-polynomial
Chapman-Enskog closure scheme (Chapman \& Cowling 1953) yields
\begin{align}\label{e3.91a}
{\bf F}&= n\,e\left(\frac{{\bf j}_\parallel}{\sigma_\parallel}
+\frac{{\bf j}_\perp}{\sigma_\perp}\right) -0.71\,n\,\nabla_\parallel T_e
-\frac{3\,n}{2\,{\mit\Omega}_e\,\tau_e}\,{\bf b}\times\nabla_\perp T_e,
\\[0.5ex]
w_i &= \frac{3\,m_e}{m_i} \frac{n\,(T_e-T_i)}{\tau_e},\\[0.5ex]
w_e &= -w_i + \frac{ {\bf j}\cdot {\bf F} }{n \,e}.
\end{align}
Here, ${\bf b}={\bf B}/B$ is a unit vector parallel to the magnetic field, and ${\bf j}= n\,e\,({\bf V}_i-{\bf V}_e)$ is the plasma current density. 
Moreover, the {\em parallel electrical conductivity}\/ is given by
\begin{equation}\label{e3.83}
\sigma_\parallel = 1.96\,\frac{n \,e^2\,\tau_e}{m_e}.
\end{equation} 
whereas the {\em perpendicular electrical conductivity}\/ takes the form
\begin{equation}\label{e3.92}
\sigma_\perp = 0.51\,\sigma_\parallel = \frac{n\,e^2\,\tau_e}{m_e}.
\end{equation}
Note that $\nabla_\parallel(\cdots) \equiv [{\bf b}\cdot\nabla
(\cdots)]\,{\bf b}$ denotes a
gradient parallel to the magnetic field, whereas $\nabla_\perp \equiv
 \nabla-\nabla_\parallel$ denotes a gradient perpendicular to the magnetic
field. Likewise, ${\bf j}_\parallel \equiv ({\bf b}\cdot{\bf j})\,{\bf b}$
represents the component of the plasma current density  flowing parallel to the
magnetic field, whereas ${\bf j}_\perp \equiv {\bf j} - {\bf j}_\parallel$
represents the perpendicular component of the plasma current density.

In a highly magnetized plasma, the electron and ion heat flux densities are written 
\begin{align}
{\bf q}_e &= -\kappa_{\parallel\,e}\,\nabla_\parallel T_e -\kappa_{\perp\,e}\,
\nabla_\perp T_e
-\kappa_{\times\,e}\,{\bf b}\times\nabla_\perp T_e- 0.71\,\frac{T_e}{e}\,{\bf j}_\parallel-
\frac{3\,T_e}{2\,{\mit\Omega}_e\,\tau_e\,e}\,{\bf b}\times{\bf j}_\perp,\label{e3.94a}\\[0.5ex]
{\bf q}_i &= -\kappa_{\parallel\,i}\,\nabla_\parallel T_i -\kappa_{\perp\,i}\,
\nabla_\perp T_i
+\kappa_{\times\,i}\,{\bf b}\times\nabla_\perp T_i,\label{e3.94b}
\end{align}
respectively. Here, the {\em parallel thermal conductivities}, which control the diffusion of heat parallel to magnetic field-lines, are 
given by
\begin{align}\label{e3.87a}
\kappa_{\parallel\,e} &= 3.2\,\,\frac{n\,\tau_e\,T_e}{m_e},\\[0.5ex]
\kappa_{\parallel\,i} &= 3.9\,\,\frac{n\,\tau_i\,T_i}{m_i},\label{e3.87b}
\end{align} 
whereas the {\em perpendicular thermal conductivities}, which control the diffusion of heat perpendicular to magnetic flux-surfaces, take the form
\begin{align}
\kappa_{\perp\,e} &= 4.7\,\frac{n\,T_e}{m_e\,{\mit\Omega}_e^{\,2}\,\tau_e},\\[0.5ex]
\kappa_{\perp\,i} &= 2\, \frac{n\,T_i}{m_i\,{\mit\Omega}_i^{\,2}\,\tau_i}.
\end{align}
Finally, the {\em cross thermal conductivities}, which control the flow of heat within magnetic flux-surfaces,  are written
\begin{align}
\kappa_{\times\,e} &= \frac{5\,n\,T_e}{2\,m_e\,{\mit\Omega}_e},\\[0.5ex]
\kappa_{\times\,i} &=\frac{5\,n\,T_i}{2\,m_i\,{\mit\Omega}_i}.
\end{align}

@@

The tensor $\bpi_0$ describes what is known as {\em parallel viscosity}.
This is a viscosity that controls the variation along magnetic field-lines of the
velocity component parallel to  field-lines.
 The parallel
viscosity coefficients are given by 
\begin{align}\label{e3.89a}
\eta_{0\,e} &= 0.73\,n\,\tau_e\,T_e,\\[0.5ex]
\eta_{0\,i} &= 0.96\,n\,\tau_i\,T_i.\label{e3.89b}
\end{align}

The tensors $\bpi_1$ and $\bpi_2$ describe what is known
as {\em perpendicular viscosity}. This is a viscosity
that controls the variation perpendicular to magnetic field-lines
of the velocity components perpendicular to field-lines. The perpendicular
viscosity coefficients are given by
\begin{align}
\eta_{1\,e} &= 0.51\, \frac{n\,T_e}{{\mit\Omega}_e^{\,2}\,\tau_e},\\[0.5ex]
\eta_{1\,i} &= \frac{3\, n\,T_i}{10\,{\mit\Omega}_i^{\,2}\,\tau_i}.
\end{align}

Finally, the tensors $\bpi_3$ and $\bpi_4$ describe what is known
as {\em gyroviscosity}. This is not really viscosity at all, because the
associated viscous stresses are always perpendicular to the velocity, implying that
there is no dissipation ({\rm i.e.}, viscous heating) associated with
this effect. The gyroviscosity coefficients are given by 
\begin{align}
\eta_{3\,e} &= -\frac{n\,T_e}{2\,{\mit\Omega}_e} ,\\[0.5ex]
\eta_{3\,i} &= \frac{n\,T_i}{2\,{\mit\Omega}_i}.
\end{align}

\subsection{Normalization of Braginskii Equations}\label{s3.9}
As we have just seen, the Braginskii equations contain terms that describe a
very wide range of different physical phenomena. For this reason, they are
extremely complicated in nature. Fortunately, however, it is not generally 
necessary to retain all of
the terms in these equations when investigating tearing mode dynamics in tokamak plasmas. In this section, we shall construct a systematic
normalization scheme for the Braginskii equations that will enable us
to determine which terms to keep, and which to discard, when investigating
tearing mode dynamics.

 It is convenient to split the friction force density,
${\bf F}$, into a component, ${\bf F}_U$, corresponding to resistivity, and a
component, ${\bf F}_T$, corresponding to the so-called {\em thermal force density}. Thus,
\begin{equation}
{\bf F} = {\bf F}_U+{\bf F}_T,
\end{equation}
where
\begin{align}
{\bf F}_U&= n\,e\left(\frac{{\bf j}_\parallel}{\sigma_\parallel}
+\frac{{\bf j}_\perp}{\sigma_\perp}\right), \\[0.5ex]
{\bf F}_T &= - 0.71\,n\,\nabla_\parallel T_e
-\frac{3\,n}{2\,{\mit\Omega}_e\,\tau_e}\,{\bf b}\times\nabla_\perp T_e.\label{e43}
\end{align}
Likewise, the electron collisional energy gain term, $w_e$, is split
into a component, $-w_i$, corresponding to the energy lost per unit volume to the ions (in the
ion rest frame), a component, $w_U$, corresponding to work done by the friction
force density, ${\bf F}_U$, and a component, $w_T$, corresponding to work done by the
thermal force density, ${\bf F}_T$. Thus,
\begin{equation}
w_e = -w_i + w_U + w_T,
\end{equation}
where
\begin{align}
w_U &= \frac{{\bf j}\cdot{\bf F}_U}{n\,e},\\[0.5ex]
w_T&= \frac{{\bf j}\cdot{\bf F}_T}{n\,e}.
\end{align}
Finally, it is helpful to split the electron heat flux density, ${\bf q}_e$, into
a diffusive component, ${\bf q}_{Te}$, and a convective component, ${\bf q}_{Ue}$. 
Thus,
\begin{equation}
{\bf q}_e = {\bf q}_{Te} + {\bf q}_{Ue},
\end{equation}
where
\begin{align}
{\bf q}_{Te} &=-\kappa_{\parallel\,e}\,\nabla_\parallel T_e -\kappa_{\perp\,e}\,
\nabla_\perp T_e
-\kappa_{\times\,e}\,{\bf b}\times\nabla_\perp T_e,\\[0.5ex]
{\bf q}_{Ue}&=  0.71\,\frac{T_e}{e}\,{\bf j}_\parallel-
\frac{3\,T_e}{2\,{\mit\Omega}_e\,\tau_e\,e}\,{\bf b}\times{\bf j}_\perp.
\end{align}

@@

The normalization procedure is designed to make all hatted quantities ${\cal O}(1)$.
The normalization of the electric field is chosen 
 such that the ${\bf E}\times{\bf B}$
velocity is of  similar magnitude to the electron diamagnetic velocity.  (See Section~\ref{lowest}.) Note that the parallel viscosity
makes an ${\cal O}(1)$ contribution to $\widehat{\bpi}_e$, whereas the gyroviscosity
makes an ${\cal O}(\zeta_e)$ contribution, and the perpendicular viscosity only 
makes an ${\cal O}(\zeta_e^{2})$ contribution. Likewise, the parallel thermal
conductivity 
makes an ${\cal O}(1)$ contribution to $\widehat{\bf q}_{Te}$, whereas the cross
conductivity 
makes an ${\cal O}(\zeta_e)$ contribution, and the perpendicular conductivity only 
makes an ${\cal O}(\zeta_e^{2})$ contribution. Similarly, the parallel components
of ${\bf F}_T$ and ${\bf q}_{Ue}$ are ${\cal O}(1)$, whereas the perpendicular
components are ${\cal O}(\zeta_e)$.

@@

As before, the normalization procedure is designed to make all hatted quantities ${\cal O}(1)$.
The normalization of the electric field is chosen 
 such that the ${\bf E}\times{\bf B}$
velocity is of similar  magnitude to the ion diamagnetic velocity. Note that the parallel viscosity
makes an ${\cal O}(1)$ contribution to $\widehat{\bpi}_i$, whereas the gyroviscosity
makes an ${\cal O}(\zeta_i)$ contribution, and the perpendicular viscosity only 
makes an ${\cal O}(\zeta_i^{2})$ contribution. Likewise, the parallel thermal
conductivity 
makes an ${\cal O}(1)$ contribution to $\widehat{\bf q}_{i}$, whereas the cross
conductivity 
makes an ${\cal O}(\zeta_i)$ contribution, and the perpendicular conductivity only 
makes an ${\cal O}(\zeta_i^{2})$ contribution. Similarly, the parallel component
of ${\bf F}_T$ is  ${\cal O}(1)$, whereas the perpendicular
component is  ${\cal O}(\zeta_i\,\mu)$.

@@

\section{Cylindrical Tearing Mode Theory}\label{cyl}
\subsection{Introduction}
The aim of this section is to describe the simplest theory of tearing mode dynamics in tokamak plasmas, according to which the plasma
equilibrium is approximated as a periodic cylinder.

@@

\subsection{Perturbed Magnetic Field}
Consider a tearing mode perturbation that has $m$ periods in the poloidal direction,
and $n$ periods in the toroidal direction, where $m/n\sim {\cal O}(1)$.  We can write the perturbed magnetic field in the divergence-free form
\begin{equation}\label{e93}
\delta{\bf B} \simeq \nabla\times (\psi\,{\bf e}_z),
\end{equation}
where
\begin{equation}\label{e93a}
\psi(r,\theta,\varphi,t)= \psi(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\varphi)].
\end{equation}
Here, $\varphi=z/R_0$ is a simulated toroidal angle. The perturbed current density becomes
\begin{equation}\label{e94}
\mu_0\,\delta {\bf j} \simeq - \nabla^2\psi\,{\bf e}_z.
\end{equation}
It is assumed that $|\delta{\bf B}|\ll |{\bf B}|$.

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Wesson 1978; Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}\simeq 0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}
